---
title: "Nanostring Analysis for Dataset X v3.0"
author: "Nathan T. Johnson"
date: "3/12/2020"
output:
  pdf_document: default
purpose: Update meta data for Nanostring to reflect what was sent, Is Nanostring comparative with Flow Cytometry?, Add CD8a to the heatmap for T-cell data for comparison,Normalize the heatmap plots so easier to compare to each gene, PD1 modulation database (Literature Search), Can a composite score for Cell Stress be measured by Nanostring data?
---


```{r results='asis', echo=FALSE, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
#load in libraries
Packages <- c('tidyverse','ggplot2','openxlsx','RColorBrewer','Hmisc','icd','hrbrthemes','viridis','knitr','PCAtools','biomaRt')
lapply(Packages, library, character.only = TRUE)
```

```{r function definitions, echo = FALSE}

#read in all sheets from excel
read_all_sheets = function(xlsxFile, ...) {
  sheet_names = openxlsx::getSheetNames(xlsxFile)
  sheet_list = as.list(rep(NA, length(sheet_names)))
  names(sheet_list) = sheet_names
  for (sn in sheet_names) {
    sheet_list[[sn]] = openxlsx::read.xlsx(xlsxFile, sheet=sn, ...)
  }
  return(sheet_list)
}

```


```{r data cleaning and setup, echo=FALSE, message=FALSE, warning=FALSE}

#set working directory
setwd('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/')


#Pull in Nanostring metadata and expression data
Expression_Data <- read.csv('Expression Data.csv')
Metadata <- read.csv('Metadata.csv')
Genes <- read.csv('Nanostring_LBL-10540-02_nCounter_NHP_Immunology_V2_Panel_Gene_List.csv')
Annotations <- read.csv('Nanostring_LBL-10540-02_nCounter_NHP_Immunology_V2_Panel_Gene_List_Annotations.csv')

#Pull in Flow Data
#The first shows immune cells as a proportion of total CD45+

CD45_flow <- read.csv('../Flow Data/percent CD45 - Nathan Modified.csv')

#The other one has sheets with data from S1 or S3, as a proportion of all cells collected or as a proportion of live cells (as indicated). Let me know if you have any questions.
all_cells_flow <- read_all_sheets('../Flow Data/immune profiling - P0003-31.xlsx')

```



# Guiding Questions 

Which immune cell type (and other) to compare with what we learned from flow cytometry?

Did PD1 elicit any changes in expression profile compared with IgG?  

What can we learn about IFNg treatment.immune cells activated?  

# What are the characteristics of the samples?

### Basic Setup

```{r,echo=FALSE}
Metadata %>% dplyr::select(-c("File.Name","Sample.Name","Cartridge.ID","Lane.Number","Import.Date"))
```



# What is being quantified?

### What Gene functions are being quantified?

```{r,echo=FALSE}
Annotations %>% colnames() %>% knitr::kable()

```



### How many genes represent each function type Information?

```{r,echo=FALSE}
Annotations %>% group_by(Cell.Type) %>% count() %>% knitr::kable()
Annotations %>% group_by(Cellular.Stress) %>% count() %>% knitr::kable()
Annotations %>% group_by(Complement.System) %>% count() %>% knitr::kable()
Annotations %>% group_by(Death.Receptor.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Extracellular.matrix.organization) %>% count() %>% knitr::kable()
Annotations %>% group_by(Fc.Receptor.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Innate.Immunity) %>% count() %>% knitr::kable()
Annotations %>% group_by(Interferon.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Interleukin.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(NF.kB) %>% count() %>% knitr::kable()
Annotations %>% group_by(MAPK.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Metabolism) %>% count() %>% knitr::kable()
Annotations %>% group_by(TLR.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(VEGF.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Wnt.Signaling) %>% count() %>% knitr::kable()
```

### What comparisons can be made?

```{r,echo=FALSE}

Metadata %>% dplyr::select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(Fraction) %>% count() %>% knitr::kable()
Metadata %>% dplyr::select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(RNA.Solution) %>% count() %>% knitr::kable()
Metadata %>% dplyr::select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(Tumor) %>% count() %>% knitr::kable()
Metadata %>% dplyr::select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(Treatment) %>% count() %>% knitr::kable()


```





# Can we group the samples?

PCA plots give an idea what similiarties are to be expected from the samples that are in it.  

PCA 
- Calculates varability across entire dataset for each gene
- Groups components 
- Using these components it allows to check for bias and grouping within samples

```{r,echo=FALSE,message=FALSE,warning=FALSE}


#expression data formatting
#remove gene controls from expression data
PCA_Data <- Expression_Data %>% filter(!grepl("NEG",Gene_Name) & !grepl("POS",Gene_Name))
row.names(PCA_Data) <- PCA_Data$Gene_Name
#drop unrelevant columns
PCA_Data <- PCA_Data %>% dplyr::select(-c("Accession","Sample.ID","Gene_Name"))
PCA_Data <- as.matrix((PCA_Data))

#metadata formatting
PCA_Metadata <- Metadata %>% dplyr::select(-c("File.Name","Sample.Name","Cartridge.ID","Lane.Number","Import.Date"))
row.names(PCA_Metadata) <- Metadata$Sample.Name


#Conduct PCA
p <- pca(PCA_Data, metadata = PCA_Metadata, removeVar = 0.1)

#Plot scree plot
output <- screeplot(p,returnPlot = TRUE)
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/PCAPlot.png')

#Plot PC1 vs PC2
  biplot(p,    
         colby = 'Tumor',
         hline = 0, vline = 0,
         legendPosition = 'right', legendLabSize = 16, legendIconSize = 8.0,
         shape = 'Treatment',
         drawConnectors = FALSE,
         title = 'PC1 vs PC2',
         subtitle = 'PC1 versus PC2')
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/Pc1vsPC2.png')
  
  
#Plot each component
pairsplot(p,
triangle = FALSE,
hline = 0, vline = 0,
pointSize = 0.8,
gridlines.major = FALSE, gridlines.minor = FALSE,
colby = 'Tumor',
title = 'Pairs plot', plotaxes = TRUE,
margingaps = unit(c(0.1, 0.1, 0.1, 0.1), 'cm'))
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/Principle Component Comparison.png')

#All Plots
eigencorplot(p,
  metavars = colnames(PCA_Metadata),
  col = c('darkblue', 'blue2', 'black', 'red2', 'darkred'),
  cexCorval = 0.7,
  colCorval = 'white',
  fontCorval = 2,
  posLab = 'bottomleft',
  rotLabX = 45,
  posColKey = 'top',
  cexLabColKey = 1.5,
  scale = TRUE,
  main = 'Principle Components Correlation',
  colFrame = 'white',
  plotRsquared = FALSE)
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/Principle Component MetaData Correlation.png')

#Which Principle Component is significantly Correlated 
eigencorplot(p,
  metavars = colnames(PCA_Metadata),
  col = c('white', 'cornsilk1', 'gold', 'forestgreen', 'darkgreen'),
  cexCorval = 0.7,
  colCorval = 'white',
  fontCorval = 2,
  posLab = 'bottomleft',
  rotLabX = 45,
  posColKey = 'top',
  cexLabColKey = 1.5,
  scale = TRUE,
  main = bquote(Principal ~ component ~ Pearson ~ r^2 ~ clinical ~ correlates),
  colFrame = 'white',
  plotRsquared = TRUE,
  corFUN = 'pearson',
  corUSE = 'pairwise.complete.obs',
  signifSymbols = c('****', '***', '**', '*', ''),
  signifCutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1))
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/Signficiant Principle Component MetaData Correlation.png')


#What genes are driving the principle component
plotloadings(p,
  rangeRetain = 0.01,
  labSize = 3.0,
  title = 'What genes are driving the principle component',
  subtitle = 'PC1, PC2, PC3, PC4, PC5',
  caption = 'Top 1% variables',
  shape = 24,
  col = c('limegreen', 'black', 'red3'),
  drawConnectors = TRUE)
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/Genes Driving Principle Compoments (1-5).png')

#What genes are driving the principle component
plotloadings(p,
  rangeRetain = 0.01,
  components = getComponents(p, c(1,3,4,9)),
  labSize = 3.0,
  title = 'What genes are driving the principle component',
  subtitle = 'Focus on Significant Components',
  shape = 24,
  col = c('limegreen', 'black', 'red3'),
  drawConnectors = TRUE)
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/Genes Driving Significant Principle Compoments.png')


#PC1 and PC3 seem to be the main areas of interest
#Plot PC1 vs PC2
  biplot(p,    
         x = 'PC1',
         y = 'PC3',
         colby = 'Tumor',
         hline = 0, vline = 0,
         legendPosition = 'right', legendLabSize = 16, legendIconSize = 8.0,
         shape = 'Treatment',
         drawConnectors = FALSE,
         title = 'Component Comparison',
         subtitle = 'PC1 versus PC3')
ggplot2::ggsave('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/Analysis/Pc1vsPC3.png')


#ID gene functions [TODO]
# mart <- useMart('ENSEMBL_MART_ENSEMBL')
# mart <- useDataset('hsapiens_gene_ensembl', mart)
# 
#   getBM(mart = mart,
#     attributes = c('affy_hg_u133a', 'ensembl_gene_id',
#       'gene_biotype', 'external_gene_name'),
#     filter = 'affy_hg_u133a',
#     values = c('215281_x_at', '214464_at', '211122_s_at', '205225_at',
#       '202037_s_at', '204540_at', '215176_x_at', '205044_at', '208650_s_at',
#       '205380_at'),
#     uniqueRows = TRUE)

```

