---
title: "Nanostring Data Characteristics Dataset X"
author: "Nathan T. Johnson"
date: "4/9/2020"
output:
  pdf_document: default
purpose: What are the general characteristics of the data and functional annotation available from Nanostring
---


```{r results='asis', echo=FALSE, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
#load in libraries
Packages <- c('tidyverse','ggplot2','openxlsx','RColorBrewer','Hmisc','icd','hrbrthemes','viridis','knitr','PCAtools','biomaRt')
lapply(Packages, library, character.only = TRUE)
```

```{r function definitions, echo = FALSE}

#read in all sheets from excel
read_all_sheets = function(xlsxFile, ...) {
  sheet_names = openxlsx::getSheetNames(xlsxFile)
  sheet_list = as.list(rep(NA, length(sheet_names)))
  names(sheet_list) = sheet_names
  for (sn in sheet_names) {
    sheet_list[[sn]] = openxlsx::read.xlsx(xlsxFile, sheet=sn, ...)
  }
  return(sheet_list)
}

```


```{r data cleaning and setup, echo=FALSE, message=FALSE, warning=FALSE}

#set working directory
setwd('C:/Users/Nathan/Dropbox/Company/Xsphera/@Data/Nanostring/')


#Pull in Nanostring metadata and expression data
Expression_Data <- read.csv('Expression Data.csv')
Metadata <- read.csv('Metadata.csv')
Genes <- read.csv('Nanostring_LBL-10540-02_nCounter_NHP_Immunology_V2_Panel_Gene_List.csv')
Annotations <- read.csv('Nanostring_LBL-10540-02_nCounter_NHP_Immunology_V2_Panel_Gene_List_Annotations.csv')

#Pull in Flow Data
#The first shows immune cells as a proportion of total CD45+

CD45_flow <- read.csv('../Flow Data/percent CD45 - Nathan Modified.csv')

#The other one has sheets with data from S1 or S3, as a proportion of all cells collected or as a proportion of live cells (as indicated). Let me know if you have any questions.
all_cells_flow <- read_all_sheets('../Flow Data/immune profiling - P0003-31.xlsx')

```

# What are the characteristics of the samples?

### Basic Setup

```{r,echo=FALSE}
Metadata %>% dplyr::select(-c("File.Name","Sample.Name","Cartridge.ID","Lane.Number","Import.Date"))
```

### What comparisons can be made?

```{r,echo=FALSE}
Metadata %>% dplyr::select(-c("File.Name","Sample.Name","Cartridge.ID","Lane.Number","Import.Date")) %>% group_by(Fraction) %>% count() %>% knitr::kable()
Metadata %>% dplyr::select(-c("File.Name","Sample.Name","Cartridge.ID","Lane.Number","Import.Date")) %>% group_by(RNA.Solution) %>% count() %>% knitr::kable()
Metadata %>% dplyr::select(-c("File.Name","Sample.Name","Cartridge.ID","Lane.Number","Import.Date")) %>% group_by(Tumor) %>% count() %>% knitr::kable()
Metadata %>% dplyr::select(-c("File.Name","Sample.Name","Cartridge.ID","Lane.Number","Import.Date")) %>% group_by(Treatment) %>% count() %>% knitr::kable()
```

# What Do The Nanostring Probes Quantify?


### What Gene functions are being quantified?

Lists the number of functions that Nanostring has indicated the genes represent

```{r,echo=FALSE}
Annotations %>% colnames() %>% knitr::kable()

```


### How many genes represent each function type Information?

+ means the number of genes that ARE labelled as being associated with a function
- means the number of genes that ARE NOT labelled as being associated with a function


```{r,echo=FALSE}
Annotations %>% group_by(Cell.Type) %>% count() %>% knitr::kable()
Annotations %>% group_by(Cellular.Stress) %>% count() %>% knitr::kable()
Annotations %>% group_by(Complement.System) %>% count() %>% knitr::kable()
Annotations %>% group_by(Death.Receptor.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Extracellular.matrix.organization) %>% count() %>% knitr::kable()
Annotations %>% group_by(Fc.Receptor.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Innate.Immunity) %>% count() %>% knitr::kable()
Annotations %>% group_by(Interferon.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Interleukin.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(NF.kB) %>% count() %>% knitr::kable()
Annotations %>% group_by(MAPK.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Metabolism) %>% count() %>% knitr::kable()
Annotations %>% group_by(TLR.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(VEGF.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Wnt.Signaling) %>% count() %>% knitr::kable()
```


