---
title: "Nanostring Analysis for Dataset X v2.0"
author: "Nathan T. Johnson"
date: "3/12/2020"
output:
  pdf_document: default
purpose: Goal to identify an RNA pattern
---


```{r load in libraries,warning=FALSE,message=FALSE,echo=FALSE}

#load in libraries
library('tidyverse')
library('ggplot2')
library("openxlsx")
library("RColorBrewer")
library("Hmisc")
library("icd")
library("hrbrthemes")
library("viridis")
library("knitr")

```

```{r function definitions, echo = FALSE}

#read in all sheets from excel
read_all_sheets = function(xlsxFile, ...) {
  sheet_names = openxlsx::getSheetNames(xlsxFile)
  sheet_list = as.list(rep(NA, length(sheet_names)))
  names(sheet_list) = sheet_names
  for (sn in sheet_names) {
    sheet_list[[sn]] = openxlsx::read.xlsx(xlsxFile, sheet=sn, ...)
  }
  return(sheet_list)
}

```

```{r data cleaning and setup, echo=FALSE, message=FALSE, warning=FALSE}

#set working directory
setwd('C:/Users/Nathan/Dropbox/Company/Xsphera/Nanostring')

#Pull in metadata and expression data
Expression_Data <- read.csv('Expression Data.csv')
Metadata <- read.csv('Metadata.csv')
Genes <- read.csv('Nanostring_LBL-10540-02_nCounter_NHP_Immunology_V2_Panel_Gene_List.csv')
Annotations <- read.csv('Nanostring_LBL-10540-02_nCounter_NHP_Immunology_V2_Panel_Gene_List_Annotations.csv')



#grab all data locations
#WARNING# adding additional files will change what data is being pulled
#WARNING# opening one of the excel files while running program will throw off the location 
#path <- './Data/'
#location <- list.files(path)

#load in data
#luminex_data is stored as two data frames contained within a list
#accessing raw or diluted is done by:
#luminex_data$`raw data`
#luminex_data$`dilution x2`
#diluation worksheet contains the ids that can be mapped to the other dataset
#luminex_data <- read_all_sheets(paste(path,location[1],sep=''))

#clinical_data <- read.xlsx(paste(path,location[2],sep=''))

#feature_data <- read.xlsx(paste(path,location[3],sep=''))

#fix excel date conversion
#for(i in 4:length(clinical_data)){
# clinical_data[,i] <- convertToDate(clinical_data[,i])
#}

#for(i in 8:9){
#  feature_data[,i] <- convertToDate(feature_data[,i])
#}

# Identify what cytokine values core is confident in
#not confident
#red_cytokines <- c("IL-6","IL-8","MCP-3","IL-10","IL-1b","IFN-g","IL-1a","IL-17a","IL-2","IL-5","IL-12p70","IL-13")
#maybe confident
#orange_cytokines <- c("MICA","MIP1-a","TNFa","TGFb3")
#definitely detected
#green_cytokines <-c("MICB","GCP-2","EGF","ANG2","MIP-2","IL-1ra","MIP-1b","FRACTALKINE","IL-2Ra","G-CSF","IL-23","FGFb","TGFb2","TGFb1")

```


# Guiding Questions 

Which immune cell type (and other) to compare with what we learned from flow cytometry?

Did PD1 elicit any changes in expression profile compared with IgG?  

What can we learn about IFNg treatment.immune cells activated?  

# What are the characteristics of the samples?

### Basic Setup

```{r,echo=FALSE}
Metadata %>% select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment")
```



# What is being quantified?

### Number of Nanostring Controls

```{r, echo=FALSE}
#Expression_Data %>% group_by(Class.Name) %>% count() %>% knitr::kable()
```

### What Gene functions are being quantified?

```{r,echo=FALSE}
Annotations %>% colnames() %>% knitr::kable()

```


### How many genes represent each function type Information?

```{r,echo=FALSE}
Annotations %>% group_by(Cell.Type) %>% count() %>% knitr::kable()
Annotations %>% group_by(Cellular.Stress) %>% count() %>% knitr::kable()
Annotations %>% group_by(Complement.System) %>% count() %>% knitr::kable()
Annotations %>% group_by(Death.Receptor.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Extracellular.matrix.organization) %>% count() %>% knitr::kable()
Annotations %>% group_by(Fc.Receptor.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Innate.Immunity) %>% count() %>% knitr::kable()
Annotations %>% group_by(Interferon.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Interleukin.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(NF.kB) %>% count() %>% knitr::kable()
Annotations %>% group_by(MAPK.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Metabolism) %>% count() %>% knitr::kable()
Annotations %>% group_by(TLR.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(VEGF.Signaling) %>% count() %>% knitr::kable()
Annotations %>% group_by(Wnt.Signaling) %>% count() %>% knitr::kable()
```

### What comparisons can be made?

```{r,echo=FALSE}

Metadata %>% select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(Fraction) %>% count() %>% knitr::kable()
Metadata %>% select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(RNA.Solution) %>% count() %>% knitr::kable()
Metadata %>% select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(Tumor) %>% count() %>% knitr::kable()
Metadata %>% select('Sample.Name','RNA.Solution',"Tumor","Fraction","Treatment") %>% group_by(Treatment) %>% count() %>% knitr::kable()


```




# Can we group the samples?

PCA plots give an idea what similiarties are to be expected from the samples that are in it.  

PCA 
- Calculates varability across entire dataset for each gene
- Groups components 
- Using these components it allows to check for bias and grouping within samples

```{r,echo=FALSE,warning=FALSE}

#drop unrelevant columns
PCA_Data <- Expression_Data %>% select (-c("Accession"))

#rotate data so genes are variables
PCA_Data <- as_tibble(cbind(nms = names(PCA_Data), t(PCA_Data)))
colnames(PCA_Data) <-PCA_Data[2,]
PCA_Data <- PCA_Data[-1, ]
PCA_Data <- PCA_Data[-1, ]
sample_names <- PCA_Data$Gene_Name
PCA_Data <- sapply(PCA_Data[,2:length(PCA_Data)],as.numeric) #make numeric
PCA_Data <- as_tibble(PCA_Data)

#summarize pca
PCA_Data.pr <- prcomp(PCA_Data, center = TRUE, scale = TRUE)
summary(PCA_Data.pr)

#make tibble for plotting with meta data
output <- tibble(Sample.Name=sample_names,x=PCA_Data.pr$x[,1],y=PCA_Data.pr$x[,2],)
output <- output %>% left_join(Metadata)
```


### Samples Separate by treatment

```{r,echo=FALSE}

#plot
test <- ggplot(output,aes(x=x,y=y,color=Tumor,shape=Treatment))+
  geom_point(aes(size=Fraction)) +
  ggtitle("PCA Plot labeling Tumor, Treatment, and Fraction") +
  xlab("PC1 (37.6%)") + ylab("PC2 (17.5%)")

ggplot2::ggsave('PCAPlot.png')

```


```{r,echo=FALSE}
ggplot(output,aes(x=x,y=y,color=Tumor,shape=Treatment))+
  geom_point(aes(size=RNA.Solution)) +
  ggtitle("PCA Plot labeling Tumor, Treatment, and RNA.Solution") +
  xlab("PC1 (37.6%)") + ylab("PC2 (17.5%)")

```

#Compare PC Components

```{r,echo=FALSE}


#PC1 = treatment effect

#PC2 = variation effect


x=PCA_Data.pr$x[,1]

y=PCA_Data.pr$x[,2]


PCA_Data.pr$rotation

nComp = 2
Xhat = PCA_Data.pr$x[,1:nComp] %*% t(PCA_Data.pr$rotation[,1:nComp])
Xhat = scale(Xhat, center = -mu, scale = FALSE)

Xhat[1,]


X = PCR_Data[,1:4]
mu = colMeans(X)

Xpca = prcomp(X)

nComp = 2
Xhat = Xpca$x[,1:nComp] %*% t(Xpca$rotation[,1:nComp])
Xhat = scale(Xhat, center = -mu, scale = FALSE)

test <- Xhat[1,]


mu<-colMeans(PCA_Data)

PCA_Data.pr$loadings

library(factoextra)
fviz_pca_var(PCA_Data.pr,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

library(factoextra)
# Eigenvalues
eig.val <- get_eigenvalue(PCA_Data.pr)
eig.val
  
# Results for Variables
res.var <- get_pca_var(PCA_Data.pr)
res.var$coord          # Coordinates
res.var$contrib        # Contributions to the PCs
res.var$cos2           # Quality of representation 

max(res.var$contrib$Dim.1)





tmp %>% order('Dim.1') %>% head()




tmp<-res.var$contrib

tmp[,1]

head(res.var$cos2)

# Results for individuals
res.ind <- get_pca_ind(PCA_Data.pr)
res.ind$coord          # Coordinates
res.ind$contrib        # Contributions to the PCs
res.ind$cos2           # Quality of representation 



fold <- fold[(fold>=5) | (fold<=-5)]



tmp <- tibble(res.var$contrib)
tmp <- res.var$contrib
tmp <- data.frame(tmp)
tmp <- tmp[order(tmp$Dim.1,decreasing=TRUE),]
PCA1 <- tmp[(tmp$Dim.1>=.33),]

row.names(PCA1)

tmp <- tmp[order(tmp$Dim.1,decreasing=TRUE),]
PCA1 <- tmp[(tmp$Dim.1>=.33),]


tmp <- tmp[order(tmp$Dim.2,decreasing=TRUE),]
PCA2 <- tmp[(tmp$Dim.2>=.33),]
row.names(PCA2)




```


# Immune Cell Type Composition?
What can we learn about IFNg treatment.immune cells activated?
Which immune cell type (and other) to compare with what we learned from flow cytometry?
PD1 elicit response?

*Missing Flow Cytometry data

```{r,echo=FALSE}
#rotate expression data so genes are columns and clean it up
data <- as_tibble(cbind(nms = names(Expression_Data), t(Expression_Data)))
colnames(data) <-data[2,]
data <- data[-1, ]
data <- data[-1, ]
data <- data[-1, ]

#Cell Types Available
Cell.Types <- unique(Annotations$Cell.Type)
  
#Select genes available for Cell Type
selected_genes <- Annotations %>% filter(Cell.Type == Cell.Types[2])

#filter genes by cell type
output <- data %>% select(selected_genes$Gene)
output$Sample.Name <- data$Gene_Name
output <- output %>% left_join(Metadata)
output <- output %>% select(-c("Sample.Name","File.Name","Cartridge.ID","Lane.Number","Import.Date")) #remove unncessary colnames


#reorganize for plotting
output <- output %>% gather(key = "Gene", value = "Expression", -RNA.Solution,-Tumor,-Fraction,-Treatment)
mode(output$Expression) <- "numeric"


#plot by normal and treatments
ggplot(output,aes(x=Gene,y = Expression, fill = Tumor)) +
   geom_violin() +
   ggtitle("Is there a relationship for Tumor?")

ggplot(output,aes(x=Gene,y = Expression, fill = RNA.Solution)) +
   geom_violin() +
   ggtitle("Is there a relationship for RNA Solution?")

ggplot(output,aes(x=Gene,y = Expression, fill = Fraction)) +
   geom_violin() +
   ggtitle("Is there a relationship for Tumor?")

ggplot(output,aes(x=Gene,y = Expression, fill = Treatment)) +
   geom_violin() +
   ggtitle("Is there a relationship for RNA Solution?")
```

### Heatmaps of Each Immune Cell Type

```{r,echo=FALSE}
#rotate expression data so genes are columns and clean it up
data <- as_tibble(cbind(nms = names(Expression_Data), t(Expression_Data)))
colnames(data) <-data[2,]
data <- data[-1, ]
data <- data[-1, ]
data <- data[-1, ]

#meta information orientated to data
samples <- data.frame(data$Gene_Name)
colnames(samples) <- "Sample.Name"
samples <- samples %>% left_join(Metadata)
rownames(samples) <- samples$Sample.Name
samples <- samples %>% select(-c("Sample.Name","File.Name","Cartridge.ID","Lane.Number","Import.Date")) #remove unncessary colnames

#function for saving
save_pheatmap_png <- function(x, filename, width=1200, height=1000, res = 150) {
  png(filename, width = width, height = height, res = res)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}


#Cell Types Available
Cell.Types <- unique(Annotations$Cell.Type)

for(i in 2:length(Cell.Types)){
  #Select genes available for Cell Type
  selected_genes <- Annotations %>% filter(Cell.Type == Cell.Types[i])
  
  if(nrow(selected_genes)>=2){
    #filter genes by cell type
    output <- data %>% select(selected_genes$Gene)
    output <- sapply(output, as.numeric)
    rownames(output) <- data$Gene_Name
    
    #plot
    my_heatmap <- pheatmap::pheatmap(output,main=Cell.Types[i],annotation_row = samples)
    #save heatmap
    save_pheatmap_png(my_heatmap, paste(as.character(Cell.Types[i]),".png",sep="")) 
  }
}

```


### Heatmaps for each Nanostring Function Annotation

```{r,echo=FALSE}
#Functions Available
Functions <- unique(colnames(Annotations))

for(i in 2:length(Functions)){
  #Select genes available for Function
  selected_genes <- Annotations %>% filter((!!rlang::sym(Functions[i])) == '+')

  if(nrow(selected_genes)>=2){
    #filter genes by cell type
    output <- data %>% select(selected_genes$Gene)
        #can't figure out why, but this selection doesn't select everything correctly
    if(i==16){
      output <- subset(output, select=-c(Gene_Name,POS_B,NEG_B,NEG_C,NEG_D,NEG_F))
    }
    output <- sapply(output, as.numeric)
    rownames(output) <- data$Gene_Name
    
    #plot
    my_heatmap <- pheatmap::pheatmap(output,main=Functions[i],annotation_row = samples)
    #save heatmap
    save_pheatmap_png(my_heatmap, paste(as.character(Functions[i]),".png",sep="")) 
  }
}

```

### Focus on Interferon gamme and IgG

```{r,echo=FALSE}
#Functions Available
Functions <- unique(colnames(Annotations))

#select Interfereon
i<-12

#Select genes available for Function
selected_genes <- Annotations %>% filter((!!rlang::sym(Functions[i])) == '+')
output <- data %>% select(selected_genes$Gene)
#can't figure out why, but this selection doesn't select everything correctly
output <- subset(output, select=-c(POS_C,NEG_G))
output <- sapply(output, as.numeric)
rownames(output) <- data$Gene_Name

#fold change of IFN gamma vs IgG
fold <- gtools::foldchange(output[6,],output[7,])

#filter for high fold change
fold <- fold[(fold>=3) | (fold<=-3)]


#plot
#make into dataframe
fold <- data.frame(fold)
fold <- data.frame(t(fold))
```


```{r,echo=FALSE}
tmp <- data %>% filter(Gene_Name == 'P0010.IFNy')
tmp <- sapply(tmp, as.numeric)
tmp2 <- data %>% filter(Gene_Name == 'P0010.IgG')
tmp2 <- sapply(tmp2, as.numeric)
fold <- gtools::foldchange(tmp,tmp2)

#filter for high fold change
fold <- fold[(fold>=5) | (fold<=-5)]

#for displaying purposes
fold <- data.frame(t(fold))
fold <- data.frame(t(fold))






```


```



