---
title: "v4"
author: "Nathan T. Johnson"
date: "4/8/2020"
output: html_document
---


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
# Flow Cytometry Data
Needs modification in order to adapt RNA amounts

```{r,echo=FALSE,message=FALSE}



#pheatmap::pheatmap(output,main=CD45_flow,annotation_row = samples)





```


# Immune Cell Type Composition?
What can we learn about IFNg treatment.immune cells activated?
Which immune cell type (and other) to compare with what we learned from flow cytometry?
PD1 elicit response?

### Heatmaps of Each Immune Cell Type


### Scaling for each gene (mean and 3 Standard Deviations)

```{r,message=FALSE,echo=FALSE}
#rotate expression data so genes are columns and clean it up
data <- as_tibble(cbind(nms = names(Expression_Data), t(Expression_Data)))
colnames(data) <-data[2,]
data <- data[-1, ]
data <- data[-1, ]
data <- data[-1, ]



#meta information orientated to data
samples <- data.frame(data$Gene_Name)
colnames(samples) <- "Sample.Name"
samples <- samples %>% left_join(Metadata)
rownames(samples) <- samples$Sample.Name
samples <- samples %>% dplyr::select(-c("Sample.Name","File.Name","Cartridge.ID","Lane.Number","Import.Date")) #remove unncessary colnames

#function for saving
save_pheatmap_png <- function(x, filename, width=1200, height=1000, res = 150) {
  png(filename, width = width, height = height, res = res)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}


#Cell Types Available
Cell.Types <- unique(Annotations$Cell.Type)

for(i in 2:length(Cell.Types)){
  #Select genes available for Cell Type
  selected_genes <- Annotations %>% filter(Cell.Type == Cell.Types[i])
  
  if(nrow(selected_genes)>=2){
    #filter genes by cell type
    #not all genes are within dataset, exclude those
    output <- data %>% dplyr::select(any_of(selected_genes$Gene)) 
    if(length(colnames(output))>1){
          output <- sapply(output, as.numeric)
          rownames(output) <- data$Gene_Name
          
          #plot
          my_heatmap <- pheatmap::pheatmap(output,main=Cell.Types[i],annotation_row = samples,scale="column")
          #save heatmap
          save_pheatmap_png(my_heatmap, paste(as.character(Cell.Types[i]),"gene_scaling.png",sep="")) 
    }
  }
}
```


###Scaling based on entire image (average and SD of 3)
for(i in 2:length(Cell.Types)){
  #Select genes available for Cell Type
  selected_genes <- Annotations %>% filter(Cell.Type == Cell.Types[i])
  
  if(nrow(selected_genes)>=2){
    #filter genes by cell type
    #not all genes are within dataset, exclude those
    output <- data %>% dplyr::select(any_of(selected_genes$Gene)) 
    if(length(colnames(output))>1){
          output <- sapply(output, as.numeric)
          rownames(output) <- data$Gene_Name
          
          #plot
          my_heatmap <- pheatmap::pheatmap(output,main=Cell.Types[i],annotation_row = samples)
          #save heatmap
          save_pheatmap_png(my_heatmap, paste(as.character(Cell.Types[i]),"entire_image_scale.png",sep="")) 
    }
  }
}




