---
title: "EPAT-Chempartner"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "U:/EPAT/Chempartnerdata")
library(tidyverse)

##Databse connection
library("RPostgreSQL")
library("odbc")
con <- DBI::dbConnect(PostgreSQL(),
                      dbname = "synthetic_lethality",
                      user    = "hbi17457",
                      password    = "bRXA6hp90UBd8o5P38Cr",
                      host = "di-comp-bio.cq8sgm5uycqj.us-east-1.rds.amazonaws.com",
                      port = 5432)
```

## EPAT Chempartner Analysis

This document is for analyzing EPAT - Chempartner data. The markdown file would do correlation of Relative IC50 values to mRNA and mutation data from CCLE cell-lines data.

## Read chempartner data and tidy them

```{r read data}
max <- readxl::read_excel("CPB-P19-22824_Summary.xlsx", sheet = "Max Inhibition", skip = 1)
min <- readxl::read_excel("CPB-P19-22824_Summary.xlsx", sheet = "Min Inhibition", skip = 1)
abs_ic50 <- readxl::read_excel("CPB-P19-22824_Summary.xlsx", sheet = "ABS", skip = 1)
rel_ic50 <- readxl::read_excel("CPB-P19-22824_Summary.xlsx", sheet = "Rel", skip = 1)
auc <- readxl::read_excel("CPB-P19-22824_Summary.xlsx", sheet = "AUC", skip = 1)
cellannot <- read.csv("cell_line_annot_master_June2020.csv", header = T) %>%
  select(name, lineage , lineage_subtype) %>%
  rename(cellline = name)

max <- max %>%
  tidyr::gather("cellline","Max_inhibition", 2:ncol(max))

min <- min %>%
  tidyr::gather("cellline","Min_inhibition", 2:ncol(min))

abs_ic50 <- abs_ic50 %>%
  tidyr::gather("cellline","abs_ic50", 2:ncol(abs_ic50))

rel_ic50 <- rel_ic50 %>%
  tidyr::gather("cellline","rel_ic50", 2:ncol(rel_ic50))

auc <- auc %>%
  tidyr::gather("cellline","auc", 2:ncol(auc))
```

## Merge all different dose response curve parameters into 1 file


```{r merge, echo= FALSE}
data <- merge(max, min , by =c("Compound/cell line" , "cellline"))
data <- merge(data , abs_ic50 , by = c("Compound/cell line" , "cellline"))
data <- merge(data , rel_ic50 , by = c("Compound/cell line" , "cellline"))
data <- merge(data , auc , by = c("Compound/cell line" , "cellline"))


data <- data %>%
  rename(compound = 'Compound/cell line') %>%
  mutate(abs_ic50 =as.numeric(gsub(">","", abs_ic50))) %>%
  mutate(rel_ic50 =as.numeric(gsub(">","", rel_ic50))) %>%
  mutate(log10rel_ic50 = log10(rel_ic50)) %>%
  mutate(cellline = gsub("-","", cellline))

data <- data %>%
  mutate(cellline = gsub("-| |/|\\.","", cellline)) %>%
  mutate(cellline = toupper(cellline))

## Merge cellline annotation file
data <- merge(data , cellannot , by = "cellline")
```


## Following chunk would fetch CCLE expression data from postgreSQL database
```{r, echo= FALSE}

CCLE <- dbGetQuery(con,
                       "SELECT depmap_id, name, gene, log2tpm FROM public.ccle_expression")

```

## Following code correlates IC50 values to mRNA expression values using both Pearson and Spearman method. All the results are exported as text file

```{r Gene expression correlation , echo= FALSE}
compound_list <- dplyr::pull(data, compound) %>%
  unique()


datalist = list()

for (i in 1:length(compound_list)) {
specificcompound <- filter(data , compound == compound_list[i])

specificcompound_with_CCLE <- inner_join(specificcompound , CCLE , by = c("cellline" = "name"))


CorrelationsOut <-
  specificcompound_with_CCLE %>%
  select(
    compound,
    cellline,
    gene,
    log2tpm,
    log10rel_ic50) %>%
  na.omit(.) %>%
  group_by(compound,gene) %>%
  summarise(
    spearman = cor.test(log2tpm,log10rel_ic50,method = "spearman" , exact = FALSE)$estimate,
    spearman.Pvalue = cor.test(log2tpm,log10rel_ic50,method = "spearman" , exact = FALSE , adjust = none)$p.value ,
    pearson = cor.test(log2tpm,log10rel_ic50,method = "pearson" , exact = FALSE)$estimate,
    pearson.Pvalue = cor.test(log2tpm,log10rel_ic50,method = "pearson" , exact = FALSE , adjust = none)$p.value ,
    df = cor.test(log2tpm,log10rel_ic50,method = "pearson" , exact = FALSE)$parameter
  )

CorrelationsOut$spearman.adjustPvalue = p.adjust(CorrelationsOut$spearman.Pvalue , method = "BH")
CorrelationsOut$pearson.adjustPvalue = p.adjust(CorrelationsOut$pearson.Pvalue , method = "BH")

datalist[[i]] <- CorrelationsOut
}

relic50_data <- dplyr::bind_rows(datalist)

write.table(relic50_data, file = "log10Rel_IC50_Correlation.txt", sep= "\t", row.names = F)
```


## Following code correlates IC50 values to mutation data using KW statistics. All the results are exported as text file
```{r}
## Mutation : Read, merge and clean files
mut <- read.table("CCLE2_mutation_matrix.txt" , sep = "\t" , header = T)

mut <- mut %>%
  gather(.,"cellline" , "Call", 2:ncol(mut)) %>%
  mutate(MutationStatus = case_when(Call == "1" ~ "Mut",
                                    TRUE ~ "WT"))

mut$cellline <- gsub("^X" , "" , mut$cellline)
mut <- rename(mut , GeneName = Hugo_Symbol)
mut$MutationStatus <- as.factor(mut$MutationStatus)


## Function to do mutation

datalist = list()


for (i in 1:length(compound_list)) {
  specificcompound <- filter(data , compound == compound_list[i])
  
  specificCompound_with_CCLEmut <-
    base::merge(
      x=mut,
      y=specificcompound,
      by = "cellline") %>%
    filter(complete.cases(.))

  ## Make a table of all the genes whose level/factor is greater than 1 - useful for KW algorithm
  genelist <- specificCompound_with_CCLEmut %>%
    filter(complete.cases(.))%>%
    select(GeneName , MutationStatus) %>%
    unique()%>%
    group_by(GeneName) %>%
    tally() %>%
    filter(n > 1)
  
  mutlist <- specificCompound_with_CCLEmut %>%
    filter(complete.cases(.))%>%
    group_by(compound,GeneName) %>%
    filter(MutationStatus == "Mut") %>%
    tally()
  
  if(nrow(genelist) != 0){  
    MutCorrelationsOut <-
      specificCompound_with_CCLEmut %>%
      filter(GeneName %in% genelist$GeneName) %>%
      select(
        cellline,
        GeneName,
        MutationStatus,
        compound,
        log10rel_ic50) %>% 
      group_by(compound,GeneName) %>%
      #dplyr::mutate(mut_n = length(MUTATIONSTATUS == "MUT")) %>%
      dplyr::summarise(
        hstat = kruskal.test(log10rel_ic50, MutationStatus , p.adjust = "bonferroni" , kw=TRUE)$statistic,
        pvalue = kruskal.test(log10rel_ic50, MutationStatus)$p.value
      )
  
    MutCorrelationsOut$adjustPvalue = p.adjust(MutCorrelationsOut$pvalue , method = "BH")
    
    summary <- specificCompound_with_CCLEmut %>%
      group_by(GeneName , compound , MutationStatus) %>%
      summarise(Mean = mean(log10rel_ic50)) %>%
      tidyr::spread(MutationStatus , Mean) %>%
      mutate(FoldChange = (Mut - WT)) %>%
      mutate(Direction = case_when(FoldChange > 0 ~ "Positive",
                                   TRUE ~ "Negative"))
    
    MutCorrelationsOut <- base::merge(
      MutCorrelationsOut,
      mutlist,
      by = "GeneName") %>%
      select(-compound.y) %>% 
      rename(compound = compound.x) %>%
      left_join(summary, by = c("compound" , "GeneName"))
  
  }
  datalist[[i]] <- MutCorrelationsOut
}

mut_data <- dplyr::bind_rows(datalist)

  
write.table(mut_data, file = "log10Rel_IC50_MutCorrelation.txt", sep= "\t", row.names = F)  
  

```

